---
title: "Creating Climate Scenarios Workflow"
author: "Fire Futures"
date: "2023-04-25"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(strex)
library(tidyverse)
library(lubridate)
library(purrr)
```

This Rmarkdown provides step by step instructions for creating climate scenarios and exporting them into RHESSys formatted time series 

#1. Cal-Adapt cleaning steps ... if you have all of the data downloaded from the cal-adapt website
```{r}
source("/capstone/firefutures/data_cleaning/cal_adapt_4models_allscenarios.R")
```

Alternatively, use CalAdaptR API or read in your own file in this format:

******
FORMAT
******

#2 Building runs 

Sourcing all of the functions in the build-runs workflow 
```{r}
# Get list of files in the directory
files <- list.files("./build_runs-functions", full.names = TRUE)

# Source all files using lapply
lapply(files, source)
```

A. Accessing unfiltered data frame from 1950-2100:
If you used the data cleaning step above & have multiple files in your global environment in this format: model_rcp#_cell# ("MIROC5_rcp45_2")

```{r}
dat <- find_df(45, "MIROC5", 2) 

head(dat)
```

B. Create your table to designate desired percentile filters for each run. Make sure the length of the list is the desired number of runs you'd like. 

If you would like to sample from all available runs, input one climate variable and designate your lower and upper percentiles to be 0 & 1
(? fix the function to allow null value ?) 

```{r, Example Percentile Table}
climate_vars_list <- list(
  data.frame(
    variable = c('wind', 'precip', 'max_temp'),
    lower = c(.1, .1, .1),
    upper = c(.99, .99, .99)
  ),
  data.frame(
    variable = c('min_temp'),
    lower = c(.55),
    upper = c(.85)
  ),
  data.frame(
    variable = c('wind', 'precip', 'max_temp'),
    lower = c(.2, .2, .6),
    upper = c(.80, .80, .80)
  ),
  data.frame(
    variable = c('wind', 'precip', 'max_temp'),
    lower = c(.25, .25, .65),
    upper = c(.75, .75, .75)
  ),
  data.frame(
    variable = c('wind'),
    lower = (0),
    upper = (1) #ALL RUNS IN SAMPLE WINDOW 
  )
)
```

Finding Runs Based on Years: Here we will use the build_runs function and specify all of the information we want...
Instead of specifying rcp, gcm and sample cell you can specify a data frame (our_gcm = dat)

Two examples: 
one specifying gcm, rcp, and sample cell to find associated data
the other specifying a data frame already in global environment/in correct format 
```{r finding runs that match climate percentiles}
run_samples <<-
  build_runs(
    rcp = 45, #45 or 85
    gcm = "MIROC5", #which of the four priority models
    sample_cell = 2, #which cal-adapt gird cell
    start_date = '04/07/2023', #first year in the time series 
    sample_window = 10, #number of years below and above the run start date we'd like to filter for
    climate_table = climate_vars_list, #this is the list we defined above
    run_type = "season") #year or season. seasons defined in the build_runs function

run_samples <<-
  build_runs(
    our_gcm = dat, #df in correct format 
    start_date = '04/07/2023', #first year in the time series 
    sample_window = 10, #number of years below and above the run start date we'd like to filter for
    climate_table = climate_vars_list, #this is the list we defined above
    run_type = "year") #year or season. seasons defined in the build_runs function

head(run_samples)
```





