---
title: "Creating Climate Scenarios Workflow"
author: "Fire Futures"
date: "2023-04-25"
output: html_document
---
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(strex)
library(tidyverse)
library(lubridate)
library(purrr)
library(here)
```

This Rmarkdown provides step by step instructions for creating climate scenarios and exporting them into RHESSys formatted time series 

Sourcing all of the functions needed to build scenarios & importing example data. All functions are in the build_runs-functions folder of the climate-scenarios repository
```{r}
dat <- read.csv(url("https://raw.githubusercontent.com/fire-futures/climate-scenarios/main/MIROC_rcp45_2")) #sourcing example cleaned data 

setwd(".../climate-scenarios/build_functions")

r_files <- list.files(pattern = "\\.R$")

for (file in r_files) {
  source(file)
}
```

A. Accessing the base unfiltered data frame from 1950-2100:

- If you used the data cleaning step above & have multiple files in your global environment in this format: model_rcp#_cell# ("MIROC5_rcp45_2") 
```{r}
# dat <- find_df(45, "MIROC5", 2)
# 
# head(dat)
```

- If you imported your data frame in the correct format:
```{r}
dat <- dat %>% mutate(water_year = ifelse(month(time) > 9, year(time) + 1, year(time)))
```

B. Create your table to designate desired percentile filters for each run. Make sure the length of the list is the desired number of runs you'd like. 

-If you would like to sample from all available runs, input one climate variable and designate your lower and upper percentiles to be 0 & 1

-Run & Notes = Input for metadeta file

-repeat_segment allows you to repeat any of the climate criteria (or a few in order) your desired amount of times

```{r, Example Percentile Table}
criteria_list <- list(
  data.frame(
    run = 'year 1 season 1',
    notes = 'Likely Scenario',
    variable = c('wind', 'precip', 'max_temp'),
    lower = c(.1, .1, .1),
    upper = c(.99, .99, .99)
  ),
  data.frame(
    run = 'year 1 season 2',
    notes = 'Average-Above Average Dry',
    variable = c('min_temp'),
    lower = c(.55),
    upper = c(.85)
  ),
  data.frame(
    run = 'year 2 season 2',
    notes = 'Average: wind, precip, max_temp',
    variable = c('wind', 'precip', 'max_temp'),
    lower = c(.25, .25, .65),
    upper = c(.75, .75, .75)
  ),
  data.frame(
    run = 'year 3 season 1',
    notes = 'No specification',
    variable = c('wind'),
    lower = (0),
    upper = (1) #ALL RUNS IN SAMPLE WINDOW 
  )
)

criteria_list <- repeat_segments(criteria_list, c(1,2,3,4), 4) #repeat 1-4 four times 
```

C. Finding Runs Based on Years: Here we will use the mind_matches function and specify all of the information we want...

Instead of specifying rcp, gcm and sample cell you can specify a data frame (base_data = dat)

Two examples: 
one specifying gcm, rcp, and sample cell to find associated data

the other specifying a data frame already in global environment/in correct format 

```{r finding runs that match climate percentiles}
ui_rcp <- 45 #rcp 45 or 85
ui_gcm <- 'MIROC5' #MIROC5, CNRM_CM5, CanESM2, HadGEM2ES 
ui_sample_cell <- 2 #Any sample cell in your environment 

#Otherwise a specify data frame in ui_base_data
ui_base_data <- dat

ui_segment_type <- 'season' #year or season
#season is defined in find_matches.R currently: wet is months 11-4 and dry 5-10

ui_start_date <- '10/07/2000' #MM/DD/YYYY

ui_climate_criteria_table <- criteria_list #Defined above 

ui_sample_window <- 20 #Even number: split above and below start date 

run_samples <<-
  find_matches(
    base_data = dat,
    start_date = ui_start_date, 
    sample_window = ui_sample_window,
    climate_criteria_table = ui_climate_criteria_table, 
    segment_type = ui_segment_type,
  )

head(run_samples)
```

Randomly Select
Takes the above list of matching options and randomly selects one of each
```{r}
randomly_select(run_samples = run_samples)
```


Build the scenarios for each grid cell, convert to time series
```{r}
ui_grid_cells <- c(1,2) # Select all grid cells
get_all_grid_cells(ui_grid_cells)
```

Get Metadata
