---
title: "Creating Climate Scenarios Workflow"
author: "Fire Futures"
date: "2023-04-25"
output: html_document
editor_options:
  markdown:
    wrap: 72
---

## This Rmarkdown provides step by step instructions for creating climate scenarios and exporting them into RHESSys formatted time series.

Link to technical documentation: (add)

Summary of approach: Here, we demonstrate a process for creating a range
of unique climate scenarios by stitching together pieces of existing
climate model projections. The segments can be defined as either years
or seasons (based on wet or dry periods). To create each segment, we
filter the climate model projections to identify seasons/years that
match specified climate characteristics of interest. If multiple
years/seasons meet the criteria, a random sample of the available
segments is selected, allowing for a variety of scenarios that fit the
same baseline criteria. This sampling and reshuffling process is
repeated several times, often with varying climate criteria per segment,
to create a number of individual time segments. The resulting segments
are then joined together to form a continuous time series with daily
values for the climate variables of interest.

All functions are thoroughly documented and exist in the
"build_functions" folder.

**Glossary:**

Sample cell: The grid cell that the user will use to query data
according to the specified climate criteria, as defined below. Since
grid cells are linked in time and space, only one grid cell can be used
to form a scenario, and all other grid cells of interest are stitched
together by the same GCM, RCP, and segment as the sample cell.

Segment type: Segment type represents the type of time frame surrounding
a segment. This project is currently configured for two segment types:
year and season.

Sample window: A sampling window represents the "window" of climate
segments before and after the segment of interest that are available for
sampling. For example, if a user builds a segment for year 2025, and
specifies a sampling window of 20 years, this means that years 2025+/-
10 years (i.e. 2015 - 2030) will be used to sample from for the segment
in 2025. Similarly, for a segment in 2026, years 2016 - 2036 will be
used for sampling.

```{r Load Packages}
knitr::opts_chunk$set(echo = TRUE)
library(strex) #string manipulation
library(tidyverse)
library(lubridate) #dates
library(purrr) #applying functions on data structures
library(here) #setting file path to wd
library(sf) #spatial map
library(leaflet) #creating a map
library(caladaptr)  # remotes::install_github("ucanr-igis/caladaptr")
library(ggplot2)
library(ggrepel) #string manipulation
library(dplyr)
```

Before obtaining the data from the API, you need to specify the spatial
extent of interest, as well as what GCM model and RCP emissions
assumption you would like to use.

```{r Base Data}
# specify your RCP, GCM, and grid cells
ui_grid_cells <- c(46662, 46780, 46781) # please see map, above, and select grid cells of interest
ui_rcp <- 85 # options: 45, 85
ui_gcm <- 'MIROC5' # options: 'HadGEM2-ES', 'CNRM-CM5', 'CanESM2', 'MIROC5'
```

Sourcing all of the functions needed to build scenarios & importing
example data. All functions are in the build_segments-functions folder
of the climate-scenarios repository

```{r Source Functions}
#obtain list of files ending in .R from build_functions
r_files <- list.files("build_functions", pattern = "\\.R$", full.names = TRUE)
#source in these files
for (file in r_files) {
  source(file)
}
```

Using the Cal Adapt API, you can examine the Cal Adapt Map to find your
grid cells of interest. Click on individual grid cells to obtain the ID.

```{r Load CalAdapt Map}
# function to load Cal Adapt grid cell map
caladapt_map()
```

Now that we've specified what our data should be, we can access the API
to obtain it.

```{r Get Data}
# specify if you will use the cal adapt data api
ui_caladapt_ind = 'Y' # mark as Y if using cal adapt data through this API
# obtain grid cell ID + latitude and longitude.
#creates the data frame with the above user inputs that the get_all_grid_cell_data() function will use to grab data for
find_grid_cell_centroid(ui_grid_cells = ui_grid_cells)
# use spatial information and grid cell ID to obtain daily time series data for each grid cell
centroids_of_interest %>% pmap(get_all_grid_cell_data) #pmap applies get_all_grid_cell_data() to each grid cell of interest
```

You can also input your own data. Just Make Sure You Follow This
Structure: - naming conventions: for each grid cell, name each as
"grid_cell_X" -

columns:

     "time"         "wind"         "min_humidity" "max_humidity" "min_temp"     "max_temp"     "precip"       "water_year"   "season"   

```{r User Data Upload}
#upload data here
```

Now create your table to designate desired percentile filters for each
segment. Make sure the length of the list is the desired number of
segments you'd like. If you would like to sample from all available
segments, input one climate variable and designate your lower and upper
percentiles to be 0 & 1.

You can add information in the segment and not specification to include
within the final metadata file for your created scenario. Repeat_segment
allows you to repeat any of the climate criteria (or a few in order)
your desired amount of times

```{r, Example Percentile Table}
criteria_list <- list(
  data.frame(
    segment_spec_id = '1',
    notes = 'Likely segment',
    variable = c('wind', 'precip', 'max_temp'),
    lower = c(.10, .10, .10),
    upper = c(.90, .90, .90)
  ),
  data.frame(
    segment_spec_id = '2',
    notes = 'Above Average Dry',
    variable = c('precip'),
    lower = c(0),
    upper = c(.45)
  ),
  data.frame(
    segment_spec_id = '3',
    notes = 'Average: wind, precip, max_temp',
    variable = c('wind', 'precip', 'max_temp'),
    lower = c(.25, .25, .25),
    upper = c(.75, .75, .75)
  ),
  data.frame(
    segment_spec_id = '4',
    notes = 'No specification',
    variable = c('wind'),
    lower = (0),
    upper = (1) # All segments in the sample window 
  )
)
criteria_list <- repeat_segments(criteria_list, c(1,2,3,4), 4) #repeat segments 1-4 four times
```

Finding segments based on years:

Here we will use the find_matches function and specify all of the
information we want. Instead of specifying rcp, gcm and sample cell you
can specify a data frame (base_data = dat).

```{r finding segments that match climate percentiles}
ui_sample_cell <- grid_cell_46662 # Choose your sample grid cell of interest. It should match one of the grid cells chosen in the spatial extent defined above. 

ui_sample_cell_metaData <- 'grid_cell_46662'# In quotes for your metadata file

ui_segment_type <- 'year' #year or season
#season is defined in find_matches.R currently: wet is months October 1st-March 31st and dry April 1st-September 30th

ui_start_date <- '2000/07/01' # YYYY/MM/DD

ui_climate_criteria_table <- criteria_list #Defined above

ui_sample_window <- 20 #Even number: split above and below start date

#find_matches uses the specified climate criteria to determine which segments in the data are appropriate to sample from based on percentile specifications
segment_samples <<-
  find_matches(
    base_data = ui_sample_cell, 
    start_date = ui_start_date,
    sample_window = ui_sample_window,
    climate_criteria_table = ui_climate_criteria_table,
    segment_type = ui_segment_type,
  )

head(segment_samples)

# Count the number of elements in each list index
number_matches <- lapply(segment_samples, function(x) length(x))
# Create a dataframe with the index and the count of years/seasons
number_matches <- data.frame(Count = unlist(number_matches))
# Print the results
print(number_matches)
```

Now you can run randomly_select, which samples from the segment_samples
available and chooses one segment at random to be stitched together to
create your final time series.

```{r randomly selecting a sample for each segment}
randomly_select(segment_samples = segment_samples) #This will only work if all segments have something to sample from
```

Now you can build the scenarios for each grid cell and convert it to
RHESSys format! All of the output exists in the all_grid_cell_output
folder in your directory.

```{r }
find_all_grid_cells() #Accesses any df from the global environment with 'grid_cell' in the title
get_all_grid_cells(df_names = ui_df_names) #loops through all grid cells in desired spatial extent and stitches the randomly selected segments to create a continous time series with daily values 
```

To make sure your desired output matches your expectations, we recommend
that you plot it.

```{r}
segments_plots(time_series = grid_cell_46662_time_series, upper_year = 2010 , lower_year = 2000)
```

Finally, run this function to generate your metadata file.

```{r}
get_metadata() #adds metadata file to the all_grid_cell_output folder in your directory 
```

If you want to create another scenario, please clear your environment
before restarting the process.
